---
title: "Pursuit Problem"
author: "Rob Carnell"
date: "June 6, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(deSolve)
```

## Problem Statement

From the [Guardian](https://www.theguardian.com/science/2019/jun/03/can-you-solve-it-dogs-in-pursuit):

Four dogs are in four corners of a square of side length 1. Each dog starts 
running towards the dog immediately anti-clockwise to it. The dogs start at the 
same time, they all run at the same speed, and at every moment each dog is 
running directly towards the neighbouring dog.

During the pursuit, the dogs will run in a spiral before they all meet in the 
centre. How far does each dog travel before the group collision?

For bonus points, can you work out how far the dogs would travel if they 
started at the three corners of an equilateral triangle of side length 1, or 
the five corners of a regular pentagon of side length 1? Would the distance be 
further or shorter than in the case wih the square?

## Starting with the Square...

Assumptions, for generality

- Square is of size $L$
- dog runs at speed (magnitude of velocity) $v = ||\vec{v}||$
- $(x,y) = (0,0)$ is the lower left corner of the square
- $(x_1,y_1)$ is the position of dog 1

We can decompose the velocity into $x$ and $y$ components.  We know that the
direction of the velocity is directly at the next dog, so the components
of velocity lie along the triangle between the dogs.

$$\vec{v} = v_x \hat{i} + v_y \hat{j}$$

$$v_x = \frac{dx_1}{dt} = v \frac{x_2-x_1}{\sqrt{(x_2-x_1)^2 + (y_2-y_1)^2}}\\
v_y = \frac{dy_1}{dt} = v \frac{y_2-y_1}{\sqrt{(x_2-x_1)^2 + (y_2-y_1)^2}}$$


Note that the paths of the four dogs are symmetric.  The $x$ position of dog 1
will match the $y$ position of dog 2.  This allows us to transform the coupled
differential equations between dogs into a set of coupled differential equations
for one dog.

$$x_1 = y_2\\
y_1 = L - x_2$$


Therefore:

$$\frac{dx_1}{dt} = v \frac{L - y_1 - x_1}{\sqrt{(L - y_1 - x_1)^2 + (x_1 - y_1)^2}}\\
\frac{dy_1}{dt} = v \frac{x_1 - y_1}{\sqrt{(L - y_1 - x_1)^2 + (x_1 - y_1)^2}}$$

Check for consistency:

At $t = 0$, $x_1 = y_1 = 0$, $\frac{dx_1}{dt} = v$ and $\frac{dy_1}{dt} = 0$

Also note where $l$ is path length, $\frac{dl}{dt} = \sqrt{\frac{dx}{dt}^2 + \frac{dy}{dt}^2} = v$

### Numerical Solution

These differential equations cannot be solved analytically in this format, so
a numerical solution can be used.

#### Simple solver

Starting with coarse time steps, we see that the dogs will not meet in the middle.

```{r bigjumps, echo=FALSE}
L <- 5 # m
f <- function(delt)
{
  v <- 1 # m/s
  d <- L # distance between them to start
  x1 <- 0
  y1 <- 0
  v1x <- v
  v1y <- 0
  t <- 0
  theta1 <- 0
  i <- 0
  
  while (d > v*delt)
  {
    i <- i + 1
    t <- t + delt
    theta1 <- theta1 + atan2(v*delt, d - v*delt)
    d <- sqrt((d - delt*v)^2 + (delt*v)^2)
    x1 <- c(x1, x1[length(x1)] + v1x*delt)
    y1 <- c(y1, y1[length(y1)] + v1y*delt)
    v1x <- v*cos(theta1)
    v1y <- v*sin(theta1)
  }
  return(list(x1 = x1, y1 = y1))
}
res <- f(1)

plot(res$x1, res$y1, type = "l", xlim = c(0, L), ylim = c(0, L),
     xlab = "x", ylab = "y")
```

Each dog travels `r sum(sqrt(diff(res$x1)^2 + diff(res$y1)^2))` units in this case.

Making the time step much smaller...

```{r smallstep, echo = FALSE}
res <- f(0.001)

plot(res$x1, res$y1, type = "l", xlim = c(0, L), ylim = c(0, L),
     xlab = "x", ylab = "y")
```

Each dog travels `r sum(sqrt(diff(res$x1)^2 + diff(res$y1)^2))` units which is very close to $L$.

#### Numeric ODE solver

Using the numeric solution for ordinary differential equations for initial 
value problems in the `deSolve` package...

```{r numeric_solve, echo = TRUE}
# diff eq system
dydt <- function(t, y, parms)
{
  L <- parms$L
  v <- parms$v
  x1 <- y[1]
  y1 <- y[2]
  denom <- sqrt((L - y1 - x1)^2 + (x1 - y1)^2)
  list(c(v*(L - y1 - x1)/denom, v*(x1 - y1)/denom))
}

# initial conditions
yini <- c(y1 = 0, y2 = 0)
N <- 100
L <- 5

# vary the velocity parameter
df5_2 <- ode(y = yini, func = dydt,
          times = seq(0, L/2, length = N), 
          parms = list(L = L, v = 2))

df5_1 <- ode(y = yini, func = dydt,
             times = seq(0, L/1, length = N), 
             parms = list(L = L, v = 1))

df5_5 <- ode(y = yini, func = dydt,
             times = seq(0, L/5, length = N), 
             parms = list(L = L, v = 5))

# the track of the dog is insensitive to the velocity
plot(df5_2[,"y1"], df5_2[,"y2"], type = "l", xlim = c(0, L), ylim = c(0, L),
     xlab = "x", ylab = "y")
lines(df5_1[,"y1"], df5_1[,"y2"])
lines(df5_5[,"y1"], df5_5[,"y2"])
```

The dogs run the same distance no matter the velocity:

```{r dist, echo=FALSE}
paste("Length 5, velocity 2: ", sum(sqrt(diff(df5_2[,"y1"])^2 + diff(df5_2[,"y2"])^2)))
paste("Length 5, velocity 1: ", sum(sqrt(diff(df5_1[,"y1"])^2 + diff(df5_1[,"y2"])^2)))
paste("Length 5, velocity 5: ", sum(sqrt(diff(df5_5[,"y1"])^2 + diff(df5_5[,"y2"])^2)))
```

Plotting the $x$ and $y$ position versus time:

```{r vtime, echo=FALSE}
plot(df5_2[,"time"], df5_2[,"y1"], type = "l", xlim = c(0, L), ylim = c(0, L), 
     xlab = "time", ylab = "coordinate x or y", col = "blue")
lines(df5_2[,"time"], df5_2[,"y2"], col = "blue")
lines(df5_1[,"time"], df5_1[,"y1"], col = "red")
lines(df5_1[,"time"], df5_1[,"y2"], col = "red")
legend("topright", legend = c("Velocity 2", "Velocity 1"), col = c("blue", "red"), lty = 1)
```

Finally, note that the rate of change in the radius from the center is constant!

```{r radius, echo=TRUE}
# radius
sqrt((df5_2[,"y1"] - L/2)^2 + (df5_2[,"y2"] - L/2)^2)[1:10]
# change in radius
diff(sqrt((df5_2[,"y1"] - L/2)^2 + (df5_2[,"y2"] - L/2)^2))[1:10]
```

## Switching to polar coordinates

$$(x,y) = (r cos(\theta) + L/2,\ r sin(\theta) + L/2)$$

$$\frac{dx}{dt} = \frac{dr}{dt} cos(\theta) - r sin(\theta)\frac{d\theta}{dt}\\
\frac{dy}{dt} = \frac{dr}{dt} sin(\theta) - r cos(\theta) \frac{d\theta}{dt}$$


Equation A:

$$\frac{dr}{dt} cos(\theta) - r sin(\theta)\frac{d\theta}{dt} = v\frac{L - rsin(\theta) - L/2 - rcos(\theta) - L/2}{\sqrt{(L - rsin(\theta) - L/2 - rcos(\theta) - L/2)^2 + (rcos(\theta) + L/2 - rsin(\theta) - L/2)}}\\
 = v \frac{-r\left(sin(\theta) + cos(\theta)\right)}{\sqrt{r^2(sin(\theta)+cos(\theta))^2 + r^2(cos(\theta) - sin(\theta))^2}}\\
  = v \frac{-\left(sin(\theta) + cos(\theta)\right)}{\sqrt{(sin(\theta)+cos(\theta))^2 + (cos(\theta) - sin(\theta))^2}} = \frac{-v}{\sqrt{2}}(sin(\theta) + cos(\theta))$$

Equation B:

$$\frac{dr}{dt} sin(\theta) - r cos(\theta) \frac{d\theta}{dt} = \frac{v}{\sqrt{2}}(cos(\theta) - sin(\theta))$$

Now, make a linear combination of Eq A and Eq B to separate variables:

$cos(\theta)$ x Eq A + $sin(\theta)$ x Eq B

$$\frac{dr}{dt}cos^2(\theta) - rsin(\theta)cos(\theta) + \frac{dr}{dt}sin^2(\theta) + rsin(\theta)cos(\theta) = -\frac{v}{\sqrt{2}}cos(\theta)(sin(\theta) + cos(\theta)) + \frac{v}{\sqrt{2}}sin(\theta)(cos(\theta) - sin(\theta))\\
\frac{dr}{dt} = -\frac{v}{\sqrt{2}}$$

Integrating to find the radius function:

$$r(t) = \int_o^t-\frac{v}{\sqrt{2}} dt + C = -\frac{v}{\sqrt{2}}t + \frac{L}{\sqrt{2}}$$

$$r(t_{end}) = 0 = -\frac{v}{\sqrt{2}}t_{end} + -\frac{L}{\sqrt{2}}\\
t_{end} = \frac{L}{v}$$

$cos(\theta)$ x Eq B - $sin(\theta)$ x Eq A

$$\frac{dr}{dt} sin(\theta)cos(\theta) - r cos^2(\theta) \frac{d\theta}{dt} - \frac{dr}{dt} sin(\theta)cos(\theta) - r sin^2(\theta)\frac{d\theta}{dt} = \frac{v}{\sqrt{2}}cos(\theta)(cos(\theta) - sin(\theta)) + \frac{v}{\sqrt{2}}sin(\theta)(cos(\theta) + sin(\theta))\\
= \frac{v}{r\sqrt{2}}$$

Integrating to find the function for $\theta$:

$$\theta(t) = \int_0^t \frac{v}{r\sqrt{2}} dt + C = \int_0^t \frac{v}{\sqrt{2}}\frac{1}{-\frac{v}{\sqrt{2}}t + \frac{L}{\sqrt{2}}} dt + C\\
= -ln(r) + C\\
\theta(0) = -3\pi/4 = -ln(L/\sqrt{2}) + C\\
\theta(t) = -ln(r) + ln(L/\sqrt{2}) - 3\pi/4$$

Finally, integrating to find the distance traveled:

$$\int_0^{t_{end}} \frac{dl}{dt} dt = \int_0^{t_{end}} v dt = v t_{end} = L$$

### Plotting

```{r polar1, echo=FALSE}
L <- 5
v <- 1
t <- seq(.Machine$double.eps, L/v - 4*.Machine$double.eps, length = 1000)
r <- -v / sqrt(2) * t + L / sqrt(2)
theta <- -log(r) + log(L / sqrt(2)) - 3*pi/4
theta_d <- theta * 180 / pi

plot(t, r, type = "l", ylim = c(0, L/sqrt(2)), xlab = "t", ylab = "r")
plot(t, theta_d, type = "l", ylim = c(-3*pi/4*180/pi, 5*pi/4*180/pi), 
     xlab = "t", ylab = "Theta (degrees)")
plot(r, theta_d, type = "l", ylim = c(-3*pi/4*180/pi, 5*pi/4*180/pi), 
     xlim = c(0, L/sqrt(2)), xlab = "radius", ylab = "Theta (degrees)")
```

```{r 4dogs}
x1 <- r * cos(theta) + L/2
y1 <- r * sin(theta) + L/2
x2 <- r * cos(theta + pi/2) + L/2
y2 <- r * sin(theta + pi/2) + L/2
x3 <- r * cos(theta + pi) + L/2
y3 <- r * sin(theta + pi) + L/2
x4 <- r * cos(theta - pi/2) + L/2
y4 <- r * sin(theta - pi/2) + L/2
plot(x1, y1, type = "l", xlim = c(0, L), ylim = c(0, L), col = "red", xlab = "x",
     ylab = "y")
lines(x2, y2, col = "blue")
lines(x3, y3, col = "green")
lines(x4, y4, col = "purple")
```

## On a triangle:

```{r tri, echo=FALSE}
dydt <- function(t, y, parms)
{
  L <- parms$L
  v <- parms$v
  x1 <- y[1]
  y1 <- y[2]
  x2 <- y[3]
  y2 <- y[4]
  x3 <- y[5]
  y3 <- y[6]
  denom12 <- sqrt((x2 - x1)^2 + (y2 - y1)^2)
  denom23 <- sqrt((x3 - x2)^2 + (y3 - y2)^2)
  denom31 <- sqrt((x1 - x3)^2 + (y1 - y3)^2)
  list(
    c(
      v*(x2 - x1)/denom12, 
      v*(y2 - y1)/denom12,
      v*(x3 - x2)/denom23,
      v*(y3 - y2)/denom23,
      v*(x1 - x3)/denom31,
      v*(y1 - y3)/denom31))
}
L <- 5
v <- 1
yini <- c(y1 = 0, y2 = 0, y3 = L, y4 = 0, y5 = L/2, y6 = sqrt(3) * L / 2)
N <- 100

df <- ode(y = yini, func = dydt,
          times = seq(0, 2*L/3/v, length = N), 
          parms = list(L = L, v = v))

df <- as.data.frame(df)
names(df) <- c("time","x1","y1","x2","y2","x3","y3")

plot(df$x1, df$y1, type = "l", xlim = c(0, L), ylim = c(0, L))
lines(df$x2, df$y2)
lines(df$x3, df$y3)
```

```{r length3}
# Length of the line
sum(sqrt(diff(df$x1)^2 + diff(df$y1)^2))
```

```{r radius3, echo=TRUE}
# radius
sqrt((df$x1 - L/2)^2 + (df$y1 - L/2/sqrt(3))^2)[1:10]
# change in radius is constant
diff(sqrt((df$x1 - L/2)^2 + (df$y1 - L/2/sqrt(3))^2))[1:10]
# theta1 is a constant distance from theta2
atan2(df$y2 - L/2/sqrt(3), df$x2 - L/2)[1:10] - atan2(df$y1 - L/2/sqrt(3), df$x1 - L/2)[1:10]
``` 

$$x = r cos(\theta) + L/2\\
y = r sin(\theta) + L/2\sqrt{3})$$

$$\frac{dx}{dt} = \frac{dr}{dt} cos(\theta) - r sin(\theta)\frac{d\theta}{dt}\\
\frac{dy}{dt} = \frac{dr}{dt} sin(\theta) - r cos(\theta) \frac{d\theta}{dt}$$


Equation A:

$$\frac{dr_1}{dt} cos(\theta_1) - r_1 sin(\theta_1)\frac{d\theta_1}{dt}\\
= v\frac{r_2cos(\theta_2) + L/2 - r_1cos(\theta_1) - L/2}{\sqrt{(r_2cos(\theta_2) + L/2 - r_1cos(\theta_1) - L/2)^2 + (r_2sin(\theta_2) + L/2\sqrt{3} - r_1sin(\theta_1) - L/2\sqrt{3})^2}}$$

Note: $r_1 = r_2$ and $\theta_2 = \theta_1 + 2\pi/3$ 

therefore

$$cos(\theta_2) = cos(\theta_1 + 2\pi/3) = cos(\theta_1)cos(2\pi/3) - sin(\theta_1)sin(2\pi/3) = - \frac{1}{2}cos(\theta_1) - \frac{\sqrt{3}}{2}sin(\theta_1)\\
sin(\theta_2) = sin(\theta_1 + 2\pi/3) = sin(\theta_1)cos(2\pi/3) + cos(\theta_1)sin(2\pi/3) = - \frac{1}{2}sin(\theta_1) + \frac{\sqrt{3}}{2}cos(\theta_1)$$

$$\frac{dr_1}{dt} cos(\theta_1) - r_1 sin(\theta_1)\frac{d\theta_1}{dt}\\
= v\frac{-\frac{3}{2}cos(\theta_1) - \frac{\sqrt{3}}{2}sin(\theta_1)}{\sqrt{(-\frac{3}{2}cos(\theta_1) - \frac{\sqrt{3}}{2}sin(\theta_1))^2 + (-\frac{3}{2}sin(\theta_1) + \frac{\sqrt{3}}{2}cos(\theta_1))^2}}\\
= \frac{v}{\sqrt{3}}\left(-\frac{3}{2}cos(\theta_1) - \frac{\sqrt{3}}{2}sin(\theta_1)\right)$$

Equation B

$$\frac{dr}{dt} sin(\theta) - r cos(\theta) \frac{d\theta}{dt}\\
= \frac{v}{\sqrt{3}}\left(-\frac{3}{2}sin(\theta) + \frac{\sqrt{3}}{2}cos(\theta)\right)$$

Separating variables:

$cos(\theta)$ x Eq A + $sin(\theta)$ x Eq B

$$\frac{dr}{dt} = \frac{v}{\sqrt{3}}\left(-\frac{3}{2}cos^2(\theta) - \frac{\sqrt{3}}{2}sin(\theta)cos(\theta) - \frac{3}{2}sin^2(\theta) + \frac{\sqrt{3}}{2}cos(\theta)sin(\theta)\right)\\
=\frac{-v\sqrt{3}}{2}$$

$cos(\theta)$ x Eq B - $sin(\theta)$ x Eq A

$$\frac{d\theta}{dt} = \frac{1}{r}\frac{v}{\sqrt{3}}\left(-\frac{3}{2}sin(\theta)cos(\theta) + \frac{\sqrt{3}}{2}cos^2(\theta) + \frac{3}{2}cos(\theta)sin(\theta) + \frac{\sqrt{3}}{2}sin^2(\theta)\right) = \frac{v}{2r}$$

$$r(t) = \int_0^t \frac{-v\sqrt{3}}{2} dt + C = \frac{-v\sqrt{3}}{2} t + C\\
r(0) = L/\sqrt{3}\\
r(t) = \frac{-v\sqrt{3}}{2} t + \frac{L}{\sqrt{3}}\\
r(t_{end}) = 0\\
t_{end} = \frac{2L}{3v}$$

$$\theta(t) = \int_0^t \frac{v}{2r} dt + C = \int_0^t \frac{v}{2(\frac{-v\sqrt{3}}{2} t + \frac{L}{\sqrt{3}})} dt + C = \frac{-1}{\sqrt{3}}ln(r) + C\\
\theta_1(0) = -5\pi/6 = \frac{-1}{\sqrt{3}}ln(\frac{L}{\sqrt{3}}) + C\\
\theta_1(t) = \frac{-1}{\sqrt{3}}ln(r) + \frac{1}{\sqrt{3}}ln(\frac{L}{\sqrt{3}}) - 5\pi/6$$

Distance

$$\int_0^{t_{end}} v dt = \frac{2}{3}L$$

### Plotting

```{r polar3, echo=FALSE}
L <- 5
v <- 1
t <- seq(.Machine$double.eps, 2*L/3/v - 4*.Machine$double.eps, length = 1000)
r <- -v * sqrt(3) / 2 * t + L / sqrt(3)
theta <- -log(r)/sqrt(3) + 1/sqrt(3)*log(L / sqrt(3)) - 5*pi/6
theta_d <- theta * 180 / pi

plot(t, r, type = "l", ylim = c(0, L/sqrt(2)), xlab = "t", ylab = "r")
plot(t, theta_d, type = "l", ylim = c(-5*pi/6*180/pi, 7*pi/6*180/pi), 
     xlab = "t", ylab = "Theta (degrees)")
plot(r, theta_d, type = "l", ylim = c(-5*pi/6*180/pi, 7*pi/6*180/pi), 
     xlim = c(0, L/sqrt(2)), xlab = "radius", ylab = "Theta (degrees)")
```


```{r 3dogs}
x1 <- r * cos(theta) + L/2
y1 <- r * sin(theta) + L/2/sqrt(3)
x2 <- r * cos(theta + 2*pi/3) + L/2
y2 <- r * sin(theta + 2*pi/3) + L/2/sqrt(3)
x3 <- r * cos(theta + 4*pi/3) + L/2
y3 <- r * sin(theta + 4*pi/3) + L/2/sqrt(3)
plot(x1, y1, type = "l", xlim = c(0, L), ylim = c(0, L), col = "red", xlab = "x",
     ylab = "y")
lines(x2, y2, col = "blue")
lines(x3, y3, col = "green")
```

## Finally, on a regular Pentagon

The center is at $x = L/2$ and $y = \frac{L}{2} tan(3\pi/10)$ and radius 
$r(0) = \frac{L}{2*cos(3\pi/10)} = \frac{L}{2sin(\pi/5)}$

Making changes for the pentagon:

$$x = r cos(\theta) + L/2\\
y = r sin(\theta) + \frac{L}{2}tan(3\pi/10)$$

$r_1 = r_2$ and $\theta_2 = \theta_1 + 2\pi/5$ 

$$cos(\theta_2) = cos(\theta_1 + 2\pi/5) = cos(\theta_1)cos(2\pi/5) - sin(\theta_1)sin(2\pi/5) = cos(\theta_1)\alpha - sin(\theta_1)\beta\\
sin(\theta_2) = sin(\theta_1 + 2\pi/5) = sin(\theta_1)cos(2\pi/5) + cos(\theta_1)sin(2\pi/5) = sin(\theta_1)\alpha + cos(\theta_1)\beta$$

Equation A:

$$\frac{dr}{dt} cos(\theta) - r sin(\theta)\frac{d\theta}{dt}\\
= v\frac{(\alpha-1)cos(\theta) - \beta sin(\theta)}{\sqrt{((\alpha-1)cos(\theta) - \beta sin(\theta))^2 + ((\alpha-1)sin(\theta) + \beta cos(\theta))^2}}\\
= \frac{v}{\sqrt{(\alpha-1)^2 + \beta^2}}\left((\alpha-1)cos(\theta) - \beta sin(\theta)\right)$$

Equation B

$$\frac{dr}{dt} sin(\theta) - r cos(\theta) \frac{d\theta}{dt}\\
= \frac{v}{\sqrt{(\alpha-1)^2 + \beta^2}}\left((\alpha-1)sin(\theta) + \beta cos(\theta)\right)$$

Separating variables:

$cos(\theta)$ x Eq A + $sin(\theta)$ x Eq B

$$\frac{dr}{dt} = \frac{v}{\sqrt{(\alpha-1)^2 + \beta^2}}\left((\alpha-1)cos^2(\theta) - \beta sin(\theta)cos(\theta) + (\alpha-1)sin^2(\theta) + \beta cos(\theta)sin(\theta)\right)\\
=\frac{v(\alpha-1)}{\sqrt{(\alpha-1)^2 + \beta^2}}$$

$cos(\theta)$ x Eq B - $sin(\theta)$ x Eq A

$$\frac{d\theta}{dt} = \frac{1}{r}\frac{v}{\sqrt{(\alpha-1)^2 + \beta^2}} \left((\alpha-1)sin(\theta)cos(\theta) + \beta cos^2(\theta) - (\alpha-1)sin(\theta) + \beta sin^2(\theta)\right)\\
= \frac{v\beta}{r\sqrt{(\alpha-1)^2 + \beta^2}}$$

$$r(t) = \int_0^t \frac{v(\alpha-1)}{\sqrt{(\alpha-1)^2 + \beta^2}} dt + C\\
= \frac{v(\alpha-1)}{\sqrt{(\alpha-1)^2 + \beta^2}} t + C\\
r(0) = \frac{L}{2*cos(3\pi/10)}\\
r(t) = \frac{v(\alpha-1)}{\sqrt{(\alpha-1)^2 + \beta^2}} t + \frac{L}{2*cos(3\pi/10)}\\
r(t_{end}) = 0\\
t_{end} = \frac{-L}{2\gamma}\frac{\sqrt{(\alpha-1)^2 + \beta^2}}{v(\alpha-1)}$$

Radius at $t = 0$: `r 1/2/cos(3*pi/10)` L
Radius at $t = t_{end}$: 0

$$\theta(t) = \int_0^t \frac{v\beta}{r\sqrt{(\alpha-1)^2 + \beta^2}} dt + C\\
= \frac{\beta}{(\alpha-1)}ln(r) + C\\
\theta_1(0) = 13\pi/10 = \frac{\beta}{(\alpha-1)}ln\left(\frac{L}{2cos(3\pi/10)}\right) + C\\
\theta_1(t) = \frac{\beta}{(\alpha-1)}ln(r) - \frac{\beta}{(\alpha-1)}ln\left(\frac{L}{2cos(3\pi/10)}\right) + 13\pi/10$$

$\theta_1$ at $t = 0$: `r sin(2*pi/5)/(cos(2*pi/5) - 1)*log(1/2/cos(3*pi/10)*5) - sin(2*pi/5)/(cos(2*pi/5) - 1)*log(5/2/cos(3*pi/10)) + 13*pi/10`

Distance

$$\int_0^{t_{end}} v dt = \frac{-L}{2\gamma}\frac{\sqrt{(\alpha-1)^2 + \beta^2}}{(\alpha-1)}$$
= `r -1/2/cos(3*pi/10)*sqrt((cos(2*pi/5) - 1)^2 + (sin(2*pi/5))^2)/(cos(2*pi/5) - 1)` L

This is equivalent to the text-book answer of $L/(1-cos(2\pi/n))$ where $n$ is
the number of sides of the polygon.

```{r polar5, echo=FALSE}
L <- 5
v <- 1
t <- seq(.Machine$double.eps, -L/2/v/cos(3*pi/10)*sqrt((cos(2*pi/5) - 1)^2 + (sin(2*pi/5))^2)/(cos(2*pi/5) - 1) - 4*.Machine$double.eps, length = 1000)
r <- v * (cos(2*pi/5) - 1) / sqrt((cos(2*pi/5) - 1)^2 + (sin(2*pi/5))^2) * t + 
  L / 2 / cos(3*pi/10)
theta <- sin(2*pi/5)/(cos(2*pi/5) - 1)*log(r) - sin(2*pi/5)/(cos(2*pi/5) - 1) *
  log(L/2/cos(3*pi/10)) + 13*pi/10
theta_d <- theta * 180 / pi

x1 <- r * cos(theta) + L/2
y1 <- r * sin(theta) + L/2*tan(3*pi/10)
x2 <- r * cos(theta + 2*pi/5) + L/2
y2 <- r * sin(theta + 2*pi/5) + L/2*tan(3*pi/10)
x3 <- r * cos(theta + 4*pi/5) + L/2
y3 <- r * sin(theta + 4*pi/5) + L/2*tan(3*pi/10)
x4 <- r * cos(theta + 6*pi/5) + L/2
y4 <- r * sin(theta + 6*pi/5) + L/2*tan(3*pi/10)
x5 <- r * cos(theta + 8*pi/5) + L/2
y5 <- r * sin(theta + 8*pi/5) + L/2*tan(3*pi/10)

plot(x1, y1, type = "l", xlim = c(-L, 2*L), ylim = c(0, 2*L), col = "red", xlab = "x",
     ylab = "y")
lines(x2, y2, col = "blue")
lines(x3, y3, col = "green")
lines(x4, y4, col = "purple")
lines(x5, y5, col = "orange")
```


