---
title: "MTH331-Wk4"
author: "Rob Carnell"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Result 17

Let $a,b,c \in \mathbb{Z}$. If $a|b$ and $c|d$, then $ac | bd$

If $a|b$ and $c|d$, then there exists $n,m \in \mathbb{Z}$ such that

$$an = b$$

$$cm = d$$

Then $bd = (an)(cm)$, and $ac | (an)(cm)$ and $ac | bd$

## Result 18

Suppose that $x,y \in \mathbb{R}$ If $x^2 + 5y = y^2 + 5x$ then $x=y$ or $x+y=5$

If $x^2 + 5y = y^2 + 5x$ then

$$x^2 - 5x = y^2 - 5y$$

$$x=y$$

Suppose $y = ax + b$

$$x^2 -5x = (ax+b)^2 - 5(ax+b)$$

$$x^2 -5x = a^2x^2 + 2abx + b^2 - 5ax -5b$$

$$a^2 = 1$$

$a = \pm 1$

$$b^2 -5b = b(b-5) = 0$$

$b=0$ or $b = 5$

$$2ab -5a = a(2b-5) = -5$$

if $b=0$, then $a=1$, therefore $y = x$

if $b=5$, then $a=-1$, therefore $y = -x + 5$ or $x + y = 5$

## Result 19

Let $a,b \in mathbb{Z}$. If $a^2(b^2-2b)$ is odd, then $a$ and $b$ are odd.

if $a^2(b^2 - 2b)$ is odd, then for $k \in \mathbb{Z}$

$$a^2(b^2 - 2b) = 2k + 1$$

$$a^2(b)(b-2) = 2k + 1$$

Suppose $a=2m+x$ and $b = 2n+y$ for $m,n \in \mathbb{Z}$ and $x,y \in {0,1}$

$$(2m+x)^2(2n+y)(2n+y-2) = (4m^2+4mx + x^2)(4n^2+2ny-4n+2ny+y^2-2y)$$

now, $u,v \in \mathbb{Z}$. $4m^2 + 4mx$ is even so set it to $2u$. And $4n^2+2ny-4n+2ny-2y$ is even so set it to $2v$

$$=(2u + x^2)(2v + y^2) = 2k + 1$$

$$4uv + 2vx^2 + 2uy^2 + x^2y^2 = 2k + 1$$

the only way for the left hand side to be odd is for $x^2y^2 = 1$ therefore both $x = 1$ and $y=1$ from the original definition of $x,y$. Therefore, both $a$ and $b$ are odd.

## Result 20

Let $x,y \in mathbb{Z}$. If $x^2(y+3)$ is even, then $x$ is even or $y$ is odd.

if $x^2(y+3)$ is even, then $x^2(y+3) = 2k$

Suppose $x=2m+a$ and $y = 2n+b$ for $m,n \in \mathbb{Z}$ and $a,b \in {0,1}$

$$x^2(y+3) = (2m+a)^2(2n+b+3) = (4m^2+4ma+a^2)(2n+b+3)$$

now, $u \in \mathbb{Z}$. $4m^2 + 4ma$ is even so set it to $2u$.

$$(2u + a^2)(2n+b+3) = 2un + 2ub + 6u + 2na^2+ba^2 + 3a^2 = 2k$$

$$ = 2(un+ub+3u+na^2) + ba^2 + 3a^2 = 2k$$

Therefore $(b+3)a^2 = 0$

Either $a=0$ or $b=-3$ corresponding to even $x$ or odd $y$

## Result 21

Suppose that $x \in \mathbb{R}$ if $x^5 + 7x^3 +5x \ge x^4+x^2+8$ then $x \ge 0$

$$x^5 + 7x^3 +5x \ge x^4+x^2+8$$

$$x(x^4+7x^2+5) \ge x^4+x^2+8$$

Note that for all $x \in \mathbb{R}$, $(x^4+7x^2+5) \ge 5$ and $x^4+x^2+8 \ge 8$

Therefore, $x \ne 0$ because $0 \ge 8$ is false

$x$ cannot be negative because a negative number times a number greater than 5 would be a negative number and it can't be greater than 8.

Therefore, $x > 0$

In fact, there is a real root to the polynomial at approximately 0.863345, and $x \ge 0.8633456924760285478$

```{r}
f <- function(x) x^5-x^4+7*x^3-x^2+5*x-8
uniroot(f, lower = 0, upper = 1, tol = 1E-12)
f(0.8633456924760285478)
```

## Result 22

Suppose that $x \in \mathbb{R}$. if $x^2 + 5x < 0$ then $x < 0$

$$x^2 + 5x < 0$$

$$x(x+5) < 0$$

$x \ge 0$ is not allowed since $x(x+5)$ is positive

$-5 < x < 0$ results in a true statement since $x(x+5)$ is negative

$x < -5$ results in a false statement since $x(x+5)$ is positive

Therefore $-5 < x < 0$

## Result 23

If $n$ is an odd integer, then $8|(n^2-1)$

if $n$ is odd then for $k \in \mathbb{Z}$, $n = 2k+1$

and $(n^2-1) = (2k+1)^2 -1 = 4k^2+4k + 1 - 1 = 4k(k+1)$

$k$ can be either even, or odd

if $k$ is even, then for $m \in \mathbb{z}$, $k = 2m$

$$4k(k+1) = 4(2m)(2m+1) = 8m(2m+1)$$

therefore $8|(n^2-1)$

if $k$ is odd, then for $m \in \mathbb{z}$, $k = 2m + 1$

$$4k(k+1) = 4(2m+1)(2m+1+1) = 8(2m+1)(m+1)$$

therefore $8|(n^2-1)$

## Result 24

Let $x,y \in \mathbb{Z}$ If $3 \nmid xy$, then $3 \nmid x$ and $3 \nmid y$

$3|xy$ iff $3|x$ or $3|y$ therefore if $3 \nmid xy$, then $3 \nmid x$ and $3 \nmid y$
