---
title: "Quiz 5"
author: "RC"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{python}
import sympy as sy
import sympy.vector as sv
```

## 1

```{r}
A <- matrix(c(1, 1, -1, 3,
              1, 2, 1, 2,
              2, 5, 4, 3,
              4, 6, 0, 10), nrow = 4, byrow=TRUE)

B <- A
B[2,] <- B[2,] - B[1,]
B[3,] <- B[3,] - 2*B[1,]
B[4,] <- B[4,] - 4*B[1,]
B[3,] <- B[3,] - 3*B[2,]
B[4,] <- B[4,] - 2*B[2,]
B[1,] <- B[1,] - B[2,]
B
```

Basis for column space `r A[,1]` and `r A[,2]`

For the null space, $x_1 = 3x_3 - 4x_4$ and $x_2 = -2x_3 + x_4$, with $x_3$ and $x_4$ as free parameters

so any multiple of [-1, -1, 1, 1] should be a solution of $Ax=0$

check

```{r}
A %*% matrix(c(-1, -1, 1, 1), ncol = 1)
A %*% matrix(5*c(-1, -1, 1, 1), ncol = 1)
```

## 2

```{r}
A <- matrix(c(-2, -7, 17, -30,
              6, 4, 0, 22,
              8, 5, 1, 28,
              13, 7, 5, 41), nrow = 4, byrow=TRUE)

B <- A
B[1,] <- B[1,] + B[2,]/2
B[2,] <- B[2,] - 6*B[1,]
B[3,] <- B[3,] - 8*B[1,]
B[4,] <- B[4,] - 13*B[1,]
B[2,] <- B[2,]/34
B[3,] <- B[3,] - 45*B[2,]
B[4,] <- B[4,] - 72*B[2,]
B[1,] <- B[1,] + 5*B[2,]
B
```

Basis for column space `r A[,1]` and `r A[,2]`

For the null space, $x_1 = -2x_3 - x_4$ and $x_2 = 3x_3 - 4x_4$, with $x_3$ and $x_4$ as free parameters

so any multiple of [-3, -1, 1, 1] should be a solution of $Ax=0$

check

```{r}
A %*% matrix(c(-3, -1, 1, 1), ncol = 1)
A %*% matrix(5*c(-3, -1, 1, 1), ncol = 1)
```

## 3

```{r}
A <- matrix(c(1, -1, 0, 3,
              -1, 1, 3, -9,
              1, -1, 1, 1,
              2, -2, -1, 8), nrow = 4, byrow=TRUE)

B <- A
B[2,] <- B[2,] + B[1,]
B[3,] <- B[3,] - B[1,]
B[4,] <- B[4,] - 2*B[1,]
B[2,] <- B[2,] - 2*B[3,]
B[3,] <- B[3,] - B[2,]
B[4,] <- B[4,] + B[2,]
B
```

Basis for column space `r A[,1]` and `r A[,3]`

For the null space, $x_1 = x_2 - 3x_4$ and $x_3 = 2x_4$, with $x_2$ and $x_4$ as free parameters

so any multiple of [-2, 1, 2, 1] should be a solution of $Ax=0$

check

```{r}
A %*% matrix(c(-2, 1, 2, 1), ncol = 1)
A %*% matrix(5*c(-2, 1, 2, 1), ncol = 1)
```

## 4

```{r}
A <- matrix(c(-6, 6, 1, -3,
              -3, 3, 7, 8,
              -4, 4, -6, -11,
              6, -6, 3, 4), nrow = 4, byrow=TRUE)

B <- A
B[1,] <- B[1,] - B[2,] - B[3,]
B[2,] <- B[2,] + 3*B[1,]
B[3,] <- B[3,] + 4*B[1,]
B[4,] <- B[4,] - 6*B[1,]
B[2,] <- B[2,] - 2*B[4,]
B[3,] <- B[3,] + 6*B[2,]
B[4,] <- B[4,] - 3*B[2,]
B[3,] <- B[3,] / -11
B[4,] <- B[4,] - 4*B[3,]
B
```

Basis for column space `r A[,1]` and `r A[,3]` and `r A[,4]`

For the null space, $x_1 = x_2$ and $x_3 = 0$ and $x_4=0$ with $x_2$ as a free parameter

so any multiple of [1, 1, 0, 0] should be a solution of $Ax=0$

check

```{r}
A %*% matrix(c(1, 1, 0, 0), ncol = 1)
A %*% matrix(5*c(1, 1, 0, 0), ncol = 1)
```
