---
title: 'Quiz #3'
author: "Rob Carnell"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1

```{r}
a <- c(4, 3, 7)
b <- c(8, 9, 10)
d <- c(-1, 2, 4)

X <- matrix(c(a, b, d), nrow = 3, ncol = 3)
X[1,] <- X[1,] - X[2,]
X[2,] <- X[2,] - 3*X[1,]
X[3,] <- X[3,] - 7*X[1,]
X[2,] <- X[2,]*17
X[3,] <- X[3,]*12
X[3,] <- X[3,] - X[2,]
X[3,] <- X[3,]/113
X[2,] <- X[2,] - 187*X[3,]
X[2,] <- X[2,]/204
X[1,] <- X[1,] + X[2,] + 3*X[3,]
X

```

Therefore the vectors are linearly independent.

Check...

```{r}
X <- matrix(c(a, b, d), nrow = 3, ncol = 3)
det(X)
```

Determinant of $X>0$, therefore the vectors are linearly independent.

## 2

```{r}
a <- c(8, -1, 2)
b <- c(0, 1, 4)
d <- c(-4, 4, 13)

X <- matrix(c(a, b, d), nrow = 3, ncol = 3)
X[1,] <- X[1,] + 7*X[2,]
X[2,] <- X[2,] + X[1,]
X[3,] <- X[3,] - 2*X[1,]
X[3,] <- X[3,]/5
X[3,] <- X[3,]*4
X[3,] <- X[3,] + X[2,]
X[1,] <- X[1,] - 7/8*X[2,]
X[2,] <- X[2,]/8
```

Therefore the vectors are linearly dependent.

Check...

```{r}
X <- matrix(c(a, b, d), nrow = 3, ncol = 3)
det(X)
```

The determinant of $X$ equals zero, therefore columns are linearly dependent.

```{r}
x1 <- c(8, 2, 3)
x2 <- c(5, 1, 3)
x3 <- c(6, 1, 0)
```

$$8a + 5b = 6$$

$$2a + b = 1$$

```{r}
X <- matrix(c(8, 5, 6, 2, 1, 1), byrow = TRUE, ncol = 3)
X[1,] <- X[1,] - 7/2*X[2,]
X[2,] <- X[2,] - 2*X[1,]
X[2,] <- X[2,] / -2
X[1,] <- X[1,] - 3/2*X[2,]
X

-0.5*x1 + 2*x2
```

Therefore if $h=4.5$, the vectors are linearly dependent

*** Are there more? ***

## 4

`r c(4, 3, 2)` is not linearly dependent on `r c(1,-1,0)`

```{r}
X <- matrix(c(1, 4, 11, -1, 3, 17, 0, 2, 8), byrow = TRUE, ncol = 3)
X

X[2,] <- X[2,] + X[1,]
X[2,] <- X[2,]/7
X[3,] <- X[3,] - 2*X[2,]
X[1,] <- X[1,] - 4*X[2,]
X

-5*c(1, -1, 0) + 4*c(4,3,2)
```

`-5*(first vector) + 4*(second vector)`

## 5

`r c(-1, 2, -3)` is not linearly dependent on `r c(5,4,3)`

```{r}
X <- matrix(c(5, -1, 6, 4, 2, -1, 3, -3, 2), byrow = TRUE, ncol = 3)
X

X[1,] <- X[1,] - X[2,]
X[2,] <- X[2,] - 4*X[1,]
X[3,] <- X[3,] - 3*X[1,]
X[3,] <- X[3,] - 6/14*X[2,]
X[2,] <- X[2,] / 14
X[3,] <- X[3,] * 14/2/-46
X[2,] <- X[2,] *14
X[2,] <- X[2,] + 29*X[3,]
X[2,] <- X[2,] / 14
X[1,] <- X[1,] + 3*X[2,]
X[1,] <- X[1,] - 7*X[3,]
zapsmall(X)

det(X)
```

Therefore the first three are linearly independent.  The fourth must be linearly dependent on the first three.


```{r}
X <- matrix(c(5, -1, 6, 8, 4, 2, -1, 0, 3, -3, 2, 1), byrow = TRUE, ncol = 4)
X

# same operations
X[1,] <- X[1,] - X[2,]
X[2,] <- X[2,] - 4*X[1,]
X[3,] <- X[3,] - 3*X[1,]
X[3,] <- X[3,] - 6/14*X[2,]
X[2,] <- X[2,] / 14
X[3,] <- X[3,] * 14/2/-46
X[2,] <- X[2,] *14
X[2,] <- X[2,] + 29*X[3,]
X[2,] <- X[2,] / 14
X[1,] <- X[1,] + 3*X[2,]
X[1,] <- X[1,] - 7*X[3,]
zapsmall(X)

zapsmall(0.0326087*c(5,4,3) + 0.6413043*c(-1,2,-3) + 1.4130435*c(6,-1,2))

```

## 6

$$Ax = b$$

```{r}
A <- matrix(c(2,0,1,-1,1,1,0,3,2,1,0,1,-3,1,2), nrow=3)
b <- matrix(c(1,2,3), ncol = 1)

D <- cbind(A, b)

D[1,] <- D[1,] - D[3,]
D[3,] <- D[3,] - D[1,]
D[3,] <- D[3,] - 3*D[2,]
D[3,] <- D[3,] / -1
D[2,] <- D[2,] * 5
D[3,] <- D[3,] * 3
D[2,] <- D[2,] - D[3,]
D[2,] <- D[2,] * 2
D[1,] <- D[1,] * 5
D[1,] <- D[1,] + D[2,]
D[1,] <- D[1,] * 3
D[3,] <- D[3,] * 2
D[1,] <- D[1,] + D[3,]
D[1,] <- D[1,]/3
D[2,] <- D[2,]/2
D[3,] <- D[3,]/6
D

# choose x_4 and x_5 = 0
A %*% matrix(c(6/5, 7/5, 1/5, 0, 0), ncol = 1)
```

$$x_4, x_5 \epsilon \textbf{R}^1$$

$$x = (6/5 - 4/5x_4 - 1/5x_5, 7/5 - 3/5x_4 - 17/5x_5, 1/5 + 1/5x_4 + 4/5x_5, x_4, x_5)$$

## 7

$$f(x, y, z) = (x + y, y - z, x + 3y+z, 3x+4y-z)$$

$$f(x, y, z) = A x$$

```{r}
A <- matrix(c(1, 1, 0, 0, 1, -1, 1, 3, 1, 3, 4, -1), byrow = TRUE, ncol = 3)
A
```

check...

```{r}
f <- function(x, y, z) c(x + y, y - z, x + 3*y+z, 3*x+4*y-z)

f(1,5,7)
A %*% matrix(c(1,5,7), ncol = 1)
```

## 8

$$f(x, y, z) = (3x + z, y - 4z, x + y+z, -x+2y+3z)$$

$$f(x, y, z) = A x$$

```{r}
A <- matrix(c(3, 0, 1, 0, 1, -4, 1, 1, 1, -1, 2, 3), byrow = TRUE, ncol = 3)
A
```

check...

```{r}
f <- function(x, y, z) c(3*x + z, y - 4*z, x + y+z, -x+2*y+3*z)

f(1,5,7)
A %*% matrix(c(1,5,7), ncol = 1)
```

## 9

$$f(x, y) = (3x + 2y, 2x+y)$$

$$f(x, y) = A x$$

```{r}
A <- matrix(c(3, 2, 2, 1), byrow = TRUE, ncol = 2)
A
```

check...

```{r}
f <- function(x, y, z) c(3*x + 2*y, 2*x+y)

f(1,9)
A %*% matrix(c(1,9), ncol = 1)
```

